% LTeX: language=en_US
\documentclass[english]{uebung_cs}
\usepackage{settings}
\blattnummer{9}
\blattname{Problem set \theblattnummer: OVH and ETH}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begin{exercise}[Fingerübungen]
  Let \( 0 < \varepsilon < 0.5 \) and \( c \in \mathbb{N} \) be arbitrary. Consider the following functions:
  \begin{align*}
    n^{2-\varepsilon},\quad
    n\cdot\log^c n,\quad
    \frac{n^2 \log\log n}{\log n},\quad
    \frac{n^2}{2^{\sqrt{n}}},\quad
    n^2,\quad
    n^{1+\varepsilon},\quad
    n^{2-\frac{1}{\log\log n}}.
  \end{align*}
  \begin{enumerate}
    \item Arrange these functions in increasing order of growth rate.
    \item Which of these functions are contained in \( n^{1+o(1)} \) and \( n^{2-o(1)} \), respectively?
  \end{enumerate}
\end{exercise}

\begin{skill}[OVH][\mandatory]
  I can use the orthogonal vectors hypothesis (OVH) to prove conditional lower bounds for polynomial-time problems.
\end{skill}

\begin{exercise}[OVH Variant]\label{ex:ovh-prime}
  Recall the Orthogonal Vectors Hypothesis (OVH):
  \begin{quote}
    \textbf{OVH}: Given two sets \( A, B \subseteq \{0, 1\}^d \) with \( |A| = |B| = n \). There is no algorithm running in time \( n^{2-\epsilon} \cdot \poly(d) \) (for any \( \epsilon > 0 \)) which decides whether there exists \( a \in A, b \in B \) such that \( a \) and \( b \) are orthogonal.
  \end{quote}
  Consider the following variant \( \text{OVH}' \) of \( \text{OVH} \):
  \begin{quote}
    \textbf{OVH}$\bm{'}$: Given a set \( A \subseteq \{0,1\}^d \) with \( |A| = n \). There is no algorithm running in time \( n^{2-\epsilon} \cdot \poly(d) \) (for any \( \epsilon > 0 \)) which decides whether there exist \( a, a' \in A \) such that \( a \) and \( a' \) are orthogonal.
  \end{quote}
  Prove that \( \text{OVH}' \) and \( \text{OVH} \) are equivalent.
\end{exercise}

\begin{exercise}[Maximum Inner Product]
  Consider the problem of finding the maximum inner product of elements of two sets:
  \begin{quote}
    \textbf{MaxInnerProduct}: Given two sets \( A, B \subseteq \mathbb{R}_{\geq 0}^d \) with \( |A| = |B| = n \), compute the maximum
    \[
      \max\{\langle a, b\rangle \mid a \in A, b \in B\},
    \]
    where \( \langle \cdot, \cdot \rangle \) denotes the standard inner product of \( \mathbb{R}_{\geq 0}^d \).
  \end{quote}
  Prove that there is no algorithm running in time \( n^{2-\epsilon} \cdot \poly(d) \) (for any \( \epsilon > 0 \)) for this problem unless OVH fails.
  \hint{You may want to use~\ref{ex:ovh-prime}.}
\end{exercise}

\begin{exercise}[Regular Expressions]
  From your algorithms classes, you may know the problem of finding a string \( P \) (often called pattern) in another string \( T \) (often called text). This well-known problem is often called Pattern Matching; there are algorithms for this problem that run in time \( \Oh(|P| + |T|) \).

  Instead of finding a single pattern string \( P \), we are now interested in finding any substring of \( T \) that can be generated by a given regular expression. Formally, consider the following problem:
  \begin{quote}
    \textbf{RegExPatternMatching}: Given a regular expression \( R \) of size \( m \), and a text \( T \) of size \( n \), determine if there is a substring \( P \) of \( T \) that can be derived from \( R \).
  \end{quote}
  Prove that there is no algorithm running in time \( \Oh((mn)^{1-\epsilon}) \) (for any \( \epsilon > 0 \)) for RegExPatternMatching unless OVH fails.
\end{exercise}

\newpage

\begin{skill}[ETH][\mandatory]
  I can use the Exponential Time Hypothesis (ETH) and the Strong Exponential Time Hypothesis (SETH) to prove conditional lower bounds for problems in polynomial time or in exponential time.
\end{skill}

\begin{exercise}[SETH$\Rightarrow$LDOVH]
  Consider the following stronger variant of the Orthogonal Vectors Hypothesis (OVH), called LDOVH (Low-dimensional OVH): For all \(\epsilon > 0\), there is a constant \(c > 0\), such that the Orthogonal Vectors problem cannot be solved in time \(\Oh(n^{2-\epsilon})\), even when the input is restricted restricted to \(d \leq c \cdot \log n\).
  Prove that SETH implies LDOVH.
  \hint{You may use the Sparsification Lemma.}
\end{exercise}

\begin{exercise}[Subsetting Sum Fun]
  A classical NP-hard problem is the Subset Sum problem:
  \begin{quote}
    \textbf{Subset Sum}: Given a set of \(n\) distinct integers \(X = \{1 \leq x_1 < \cdots < x_n\}\) and an integer \(t\), determine whether there is a subset \(A \subseteq X\) that sums up to \(t\), that is \(\sum_{a \in A} a = t\).
  \end{quote}
  Assuming the Exponential Time Hypothesis (ETH), prove that there is a \( \delta > 0 \), such that no algorithm for Subset Sum has a running time of \( \Oh^*(2^{\delta n}) \).
\end{exercise}
\end{document}
