% LTeX: language=en_US
\documentclass[english]{uebung_cs}
\usepackage{settings}
\blattnummer{8}
\blattname{Problem set \theblattnummer: Fixed-Parameter Intractability}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\textbf{Overview:} With this problem set, you can train proving \textnormal{W[1]}-hardness.

\textbf{Instructions:} For each skill, select \textbf{exactly one} problem below and submit your solution in \href{https://moodle.studiumdigitale.uni-frankfurt.de/moodle/course/view.php?id=6259}{Moodle}; in your submission, make sure to repeat the problem that you are solving.
The problems are roughly ordered by difficulty, choose a problem that you find non-trivial and interesting. (You are of course welcome to try the other problems as well and ask us for feedback.)

\begin{skill}[Falsify reduction][\mandatory]
  I can explain why a given reduction does not establish \textnormal{W[1]}-hardness.
\end{skill}

% Cygan 13.1
\begin{exercise}[Vertex-Cover]
  Let $G$ be a graph and $S\subseteq V(G)$.
  Prove that $S$ is a maximum independent set of $G$ if and only if $V(G)\setminus S$ is a minimum vertex-cover of $G$.
  Use this to give a polynomial-time reduction from Independent Set to Vertex-Cover.
  Why is this not also a parameterized reduction that establishes that Vertex-Cover is \textnormal{W[1]}-hard? Is there a parameterized reduction from Vertex-Cover to Independent Set?
\end{exercise}

\begin{skill}[Design a reduction][\mandatory]
  I can establish \textnormal{W[1]}-hardness by designing a suitable parameterized reduction from a problem known to be \textnormal{W[1]}-hard.
\end{skill}

% Cygan 13.2
\begin{exercise}[Dominating Set]
  Give a parameterized reduction from Set Cover to Dominating Set.
\end{exercise}

% Cygan 13.3
\begin{exercise}[Multicolored Biclique]
  In the Multicolored Biclique problem, the input consists of a bipartite graph~\( G \) with two parts \( A\) and~\(B \), an integer \( k \), a partition of \( A \) into \( k \) sets \( A_1, A_2, \ldots, A_k \), and a partition of \( B \) into~\( k \) sets \( B_1, B_2, \ldots, B_k \); the question is whether there exists a subgraph of \( G \) isomorphic to the biclique~\( K_{k,k} \), with one vertex in each of the sets \( A_i \) and \( B_i \). Prove that Multicolored Biclique is \textnormal{W[1]}-hard.
\end{exercise}

% Cygan 13.23
\begin{exercise}[Set Packing]
  In the Set Packing problem, the input consists of a family \( \mathcal{F} \) of subsets of a finite universe~\( U \) and an integer \( k \), and the question is whether one can find \( k \) pairwise disjoint sets in \( \mathcal{F} \). Prove that Set Packing is \textnormal{W[1]}-hard when parameterized by \( k \).
\end{exercise}

% \begin{exercise}[Isolation Lemma]
%   Prove or disprove:
%   The Isolation Lemma (Lemma 11.5) still holds if one replaces the sum in the definition of $\textbf{w}$ by a product,
%   and the $\min$ by $\max$.
% \end{exercise}

% \begin{exercise}[todo: skill]
%   Show that $\textsc{Edge Multiway Cut}$ is polynomial-time solvable on trees.
% \end{exercise}

% \begin{exercise}[todo: skill]
%   %[\easy]
%   $\textsc{Shortest Path}$ can be solved in time $O(m\cdot \log n)$ by Dijkstra's Algorithm on connected graphs without negative edge-weights. It can be solved in time $O(m \cdot n)$ by the Bellmann-Ford-Algorithm on connected graphs that may contain negative edge-weights. Show that it can be solved in time $O(m\cdot \operatorname{poly}(k,\log n))$ where $k$ is the number of edges with negative weights. Recall that the algorithm must either compute the length of a shortest path from a given node $s$ to a given node $t$ or correctly report the existence of a negative-weight cycle.
% \end{exercise}

\end{document}
